---
title: "United States COVID-19 Analysis"
author: "David J Jackson"
date:  "June 7,020 18:25 UD/UT"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=9, fig.height=6)
library(tidyverse)
library(plotly)
library(RSQLite)
library(patchwork)
library(ggeasy)
library(scales)
theme_set(theme_light())
```
```{r,echo=FALSE}
rm(list=ls())
#these libraries are necessary
#create the URL where the dataset is stored with automatic upDates every day
db <- dbConnect(SQLite(), dbname="../COVIDDB/CORVID.sqlite3")
```
```{r,echo=FALSE}
US <- dbGetQuery(db, "SELECT Countries,Reported,Cases,Deaths FROM ECDC 
                 WHERE Countries ='USA'")
US$Reported <- as.Date(US$Reported)

```

## US Cases  and Deaths by Date (Previous 60 Days)
 * Data Source: European Centre for Disease Prevention and Control
 * URL: https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide

```{r,echo=FALSE}
plot_1 <-US %>% filter(Reported >= Sys.Date() - 60) %>%  ggplot() +
  geom_col(aes(x=Reported,y=Cases,fill=Cases)) +
  theme(axis.text.x = element_text(angle = 45)) +
      labs(title="US COVID-19: Daily Cases by Date",y="Total Cases",subtitle = "(Shaded Box = incubation period)") +
  geom_rect(data=US,aes(xmin=max(Reported) -10,xmax=max(Reported) -15,ymin=min(Cases),ymax=max(Cases)), alpha= 0.01,fill="grey") 
  
ggplotly(plot_1)
```
```{r}
    
# ggplotly(cases_1)

plot_2 <-  US %>% filter(Reported >= Sys.Date() - 60) %>%  ggplot() + geom_col(aes(x=Reported,y=Deaths,fill=Deaths)) +
  theme(axis.text.x = element_text(angle = 45)) +
      labs(x="US Date Reported",y="Total Deaths",title=" US COVID-19: Deaths by Date") 
  
           
ggplotly(plot_2)

```

### US Accumulated Daily Total: Cases and Deaths
* Data Source: John Hopkins Coronavirus Resource Center
* URL: https://coronavirus.jhu.edu/map.html

#### Exponential Growth by the Numbers:
* The first confirmed case  US were on Jan. 1, 2020 (Washington State)
* March 11, 2020 total cases: 1,281
* March 17, 2020 total cases: 6,421
* March 19, 2020 total cases: 13,747
* March 21, 2020 total cases: 25,600
* April 11, 2020 total cases: 52,639

```{r,echo=FALSE}
JHU_US<- dbGetQuery(db, "SELECT country,Date,confirmed,deaths,recovered FROM JHUDATA
                     WHERE country='US';")
colnames(JHU_US) <- c("Countries","Date","Cases","Deaths","Recovered")
JHU_US$Date <- as.Date(JHU_US$Date) 
```

```{r,echo=FALSE}
JHU_US %>% filter(Date >= Sys.Date() -60) %>%
  ggplot() + geom_col(aes(x=Date,y=na.omit(Cases),fill=Cases)) +
  theme(axis.text.x = element_text(angle = 45)) +
      labs(title="USA COVID-19: Accumulated Total (Confirmed) Cases by Date",x="Date Reported",y="Total Cases")  + 
  scale_fill_gradient(labels = scales::comma)

# scale_y_continuous(labels = scales::comma)
```
```{r,echo=FALSE}

acc_4 <- JHU_US %>% filter(Date >= Sys.Date() -60) %>%
    ggplot() + geom_col(aes(x=Date,y=Deaths,fill=Deaths)) +
  theme(axis.text.x = element_text(angle = 45)) +
      labs(title="US COVID-19: Total  Deaths by Date",x="Date Reported",y="Total Deaths") +   
  scale_y_continuous(labels = scales::comma) + scale_fill_gradient(labels = scales::comma)
ggplotly(acc_4)
```

### US Recoveries Accoumliated  and Daily Totals

```{r} 
JHU_US %>% filter(Date >= Sys.Date() -60) %>%
  ggplot() + geom_col(aes(x=Date,y=na.omit(Recovered),fill=Recovered)) +
  theme(axis.text.x = element_text(angle = 45)) +
      labs(title="USA COVID-19: Accumulated  Recoveries by Date",x="Date Reported",y="Total Cases")  + scale_y_continuous(labels = scales::comma) +
  scale_fill_gradient(labels = scales::comma)
```
```{r}
US %>% ggplot(aes(x=Reported,y=Cases)) + geom_line() +
  geom_smooth() + scale_y_log10() +
  labs(title="US Accumulated Cases by Date(Log10)",x="Date",y=" Log of Cases")
```
```{r}
US %>% ggplot(aes(x=Reported,y=Deaths)) + geom_line() +
  geom_smooth() + scale_y_log10() +
  labs(title="US Accumulated Deaths by Date(Log10)",x="Date",y=" Log of Cases")
```






