---
title: 'GGPLOT2: Us in a Nutshell'
author: "David Jackson"
date: "10/11/2020"
output: 
  html_document: 
    fig_width: 10
    fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
library(RSQLite)
library(ggplot2)
library(scales)
library(plotly)
library(esquisse)
theme_set(theme_linedraw())
```

```{r,echo=FALSE}
rm(list=ls())
source("../DATA/movavg.R")
db <- db <- dbConnect(RSQLite::SQLite(),dbname= "../COVIDDB/COVID.sqlite3")
df <- dbGetQuery(db,"select * from JHUDATA")
df$date <- as.Date(df$date)
df <- df[order(df$date,decreasing = TRUE),]
df <- subset(df,country =="US")
str(df)

```

### Plots of Cumulative Cases and Deaths
#### Source: New York Times

```{r,echo=FALSE}
cumm_cases <-ggplot(df) +geom_line(aes(x=date,y=confirmed),lwd=3,col="red") + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Cumulative Total Cases",
       x="Date Reported",y="Cumulative Cases") 
```
```{r}
cumm_deaths <-ggplot(df) +geom_line(aes(x=date,y=deaths),lwd=3,col="blue") + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Cumulative Total Deaths",x="Date Reported",y="Cumulative Deaths")
ggplotly(cumm_cases)
ggplotly(cumm_deaths)
```

## GGPLOt2 Plots of Daily Deaths and Cases
#### Source: European CDPC

```{r, echo=FALSE}
df1 <- dbGetQuery(db,"select * from ECDC")
df1$Reported <- lubridate::ymd(df1$Reported)
#df1$Reported <- as.Date(df1$Reported,format="%Y-%m-%d")
df1 <- df1[order(df1$Reported,decreasing = TRUE),]
df1 <- subset(df1,Countries =="USA" )
df1$Reported <- as.Date(df1$Reported)
df1death <- df1 %>% filter(Reported>="2020-02-29") 
```

#### Average US Covid-19 Deaths/day

```{r}
mean((df1death$Deaths))
```

### Washington University
#### Institute for Health Metrics and Evaluation
* "Daily deaths is the best indicator of the progression of the pandemic, although there is generally a 17-21 day lag between infection and deaths.+
Daily deaths is the best indicator of the progression of the pandemic, although there is generally a 17-21 day lag between infection and deaths.
* Revised US COVID19 Deaths: 363,269 by Jan. 1, 2021 (Updated Oct. 2, 2020)
* Revised US COVID19 Deaths: 363,269 by Feb. 1, 2021 (Updated Oct. 9, 2020)
* Revised US COVID19 Deaths: 394,693 by Feb. 1, 2021 (Updated Oct. 10, 2021)

```{r}
##
## 363, 000 Deaths by Jan. 2010 
(363269 - sum(df1death$Deaths)) / 83  # Jan 1, 2021
sum(df1death$Deaths) + (mean((df1death$Deaths)) * 87)
##
## 363269 by Feb. 1, 2020
(363269 - sum(df1death$Deaths)) / 114  # Jan 1, 2021
sum(df1death$Deaths) + (mean((df1death$Deaths)) * 114)
##
## 394,693 by Feb. 1, 2021
##
(394693 - sum(df1death$Deaths)) / 114 # Feb 1, 2021
sum(df1death$Deaths) + (mean((df1death$Deaths)) * 114)
##
## 
```
### US COVID19
```{r,echo=FALSE}
ggplot(df1death) +geom_line(aes(x=Reported,y=Deaths),lwd=2,col="blue") + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Daily Total Deaths(2020-03-01 - 2020-10-11)",
       subtitle="Daily Avg (Red:949); Jan. 1 (Green:1794/363269);Feb.1 (Blue:1581/394693);Feb 1(Black: 1306/363269))"
       ,x="Date Reported",y="Daily Deaths") +
  geom_hline(yintercept=949,lwd=2,col="red") +
  geom_hline(yintercept=1794,lwd=1,col="green") +
  geom_hline(yintercept=1581,lwd=1,col="blue") +
  geom_hline(yintercept=1306,lwd=1,col="black")

  ```


