---
title: 'GGPLOT2: Us in a Nutshell'
author: "David Jackson"
date: "10/17/2020"
output: 
  html_document: 
    fig_width: 10
    fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
library(RSQLite)
library(ggplot2)
library(scales)
library(plotly)
library(esquisse)
theme_set(theme_linedraw())
```

```{r,echo=FALSE}
rm(list=ls())
source("../DATA/movavg.R")
db <- db <- dbConnect(RSQLite::SQLite(),dbname= "../COVIDDB/COVID.sqlite3")
df <- dbGetQuery(db,"select * from JHUDATA")
df$date <- as.Date(df$date)
df <- df[order(df$date,decreasing = TRUE),]
df <- subset(df,country =="US")
str(df)

```

### Plots of Cumulative Cases and Deaths
#### Source: New York Times

```{r,echo=FALSE}
cumm_cases <-ggplot(df) +geom_line(aes(x=date,y=confirmed),lwd=3,col="red") + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Cumulative Total Cases",
       x="Date Reported",y="Cumulative Cases") 
```
```{r}
cumm_deaths <-ggplot(df) +geom_line(aes(x=date,y=deaths),lwd=3,col="blue") + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Cumulative Total Deaths",x="Date Reported",y="Cumulative Deaths")
ggplotly(cumm_cases)
ggplotly(cumm_deaths)
```

## GGPLOt2 Plots of Daily Deaths and Cases
#### Source: European CDPC

```{r, echo=FALSE}
df1 <- dbGetQuery(db,"select * from ECDC")
df1$Reported <- lubridate::ymd(df1$Reported)
#df1$Reported <- as.Date(df1$Reported,format="%Y-%m-%d")
df1 <- df1[order(df1$Reported,decreasing = TRUE),]
df1 <- subset(df1,Countries =="USA" )
df1$Reported <- as.Date(df1$Reported)
df1death <- df1 %>% filter(Reported>="2020-02-29") 
```

#### Average US Covid-19 Deaths/day

```{r}
mean((df1death$Deaths))
```

### Washington University
#### Institute for Health Metrics and Evaluation
* "Daily deaths is the best indicator of the progression of the pandemic, although there is generally a 17-21 day lag between infection and deaths.

* Revised US COVID19 Deaths: 363,269 by Jan. 1, 2021 (Updated Oct. 2, 2020)
* Revised US COVID19 Deaths: 363,269 by Feb. 1, 2021 (Updated Oct. 9, 2020)
* Revised US COVID19 Deaths: 394,693 by Feb. 1, 2021 (Updated Oct. 10, 2020)
* Revised US COVID19 Deaths: 389,087 by Feb. 1, 2021 (Updated Oct. 15,2020)

#### US COVID-19 Deaths  Guesstimate as of Oct. 17, 2020
* Oct. 17, 2020 Fixed math error for average US Deaths per day
* Dec. 15, 2020 WU IHME Model.
* Nov 1, 2020 = 233,706
* Dec. 1, 2020 =261,895
* Jan. 1, 2021 = 291151

### Daily US COID-19 Cases

```{r}
ggplot(df1death) +geom_line(aes(x=Reported,y=Cases),col="red") + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) + 
  geom_smooth(aes(x=Reported,y=Cases),span=0.25) +
  labs(title="US Daily COVID19 Cases")
```

```{r}
##
## 363269 by Jan 1 . 2021 
(363269 - sum(df1death$Deaths)) / 77  # Jan 1, 2021
sum(df1death$Deaths) + (mean((df1death$Deaths)) * 77)  ## Jan. 1, 2021(363269)
##
##  by Feb. 1, 2020
(363269 - sum(df1death$Deaths)) / 108  # Feb 1, 2021
sum(df1death$Deaths) + (mean((df1death$Deaths)) * 108) ## Feb., 2021(363269)
##
## 394,693 by Feb. 1, 2021
##
(389087 - sum(df1death$Deaths)) / 108 # Feb 1, 2021
sum(df1death$Deaths) + (mean((df1death$Deaths)) * 108) ## Feb.1 , 2021(389087)
##
## 
```

### US COVID19 Projected Deaths

```{r,echo=FALSE}
ggplot(df1death) +geom_line(aes(x=Reported,y=Deaths),lwd=2,col="blue") + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Daily Total Deaths(2020-03-01 - 2020-10-17)",
       subtitle="Daily Avg (Red:942); Jan. 1 (Green:1878/363269);Feb.1 (Blue:1340/363269);Feb 1(Black: 1579/389087))"
       ,x="Date Reported",y="Daily Deaths") +
  geom_hline(yintercept=942,lwd=2,col="red") +
  geom_hline(yintercept=1878,lwd=1,col="green") +
  geom_hline(yintercept=1340,lwd=1,col="blue") +
  geom_hline(yintercept=1579,lwd=1,col="black")

```
### US COVID19 Projected Deaths:Loess/Lowess Smoothed

```{r,echo=FALSE}
ggplot(df1death) +geom_smooth(aes(x=Reported,y=Deaths,col="Loess"),span=0.25) +
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Daily Total Deaths(2020-03-01 - 2020-10-17)",
       subtitle="Daily Avg (Red:942); Jan. 1 (Green:1878/363269);Feb.1 (Blue:1340/389087);Feb 1(Black: 1579/389087))"
       ,x="Date Reported",y="Daily Deaths") +
  geom_hline(yintercept=942,lwd=1.5,col="red") +
  geom_hline(yintercept=1878,lwd=1.5,col="green") +
  geom_hline(yintercept=1340,lwd=1.5,col="blue") +
  geom_hline(yintercept=1579,lwd=1.5,col="black")

```
### Just COVID19 Deaths for Feb. 1, 2020 


```{r,echo=FALSE}
ggplot(df1death) +geom_line(aes(x=Reported,y=Deaths),lwd=2,col="blue") + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Daily Total Deaths(2020-03-01 - 2020-10-17)",
       subtitle="Daily Avg (Red:942);Feb 1(Black: 1579/389087))"
       ,x="Date Reported",y="Daily Deaths") +
  geom_hline(yintercept=942,lwd=1.5,col="red") +
    geom_hline(yintercept=1579,lwd=1.5,col="green")

```
