---
title: "US COVID19 Source Data Comparison"
author: "David Jackson"
date: "11/08/2020"
output: 
  html_document: 
    fig_width: 9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
library(RSQLite)
library(ggplot2)
library(scales)
library(plotly)
library(pracma)
```

#### European CDPC Mean Cases and Deaths

```{r}
rm(list=ls())
source("../DATA/movavg.R")
db <- db <- dbConnect(RSQLite::SQLite(),dbname= "../COVIDDB/COVID.sqlite3")
df <- dbGetQuery(db,"select * from ECDC")
df <- subset(df,Countries =="USA"  )
#
df <- df[order(df$Reported),]
#
df$Reported <- as.Date(df$Reported)
AA <- subset(df,Reported >="2020-01-21")
mean(AA$Cases) 
#
AB <- subset(df,Reported >="2020-02-28")
mean(AB$Deaths)
#
US <- subset(df,Reported >="2020-03-01")

US$Rate <- US$Deaths/US$Cases
```

#### New York Times Mean Cases and  Deaths

```{r}
## New York Times Data

NYT <- dbGetQuery(db,"select * from STATESDAILY")
NYT$date <- as.Date(NYT$date)
NYTC <- subset(NYT,date >="2020-01-21")
NYTCASES <-aggregate(new_cases~date,NYTC,FUN=sum)

NYTD <- subset(NYT,date >="2020-02-28")
NYTDEATHS <-aggregate(new_deaths~date,NYTD,FUN=sum)
## Mean Cases and Deaths
##
mean(NYTCASES$new_cases)
mean(NYTDEATHS$new_deaths)

```

#### John Hopkins Mean Cases and Deaths

```{r}
##
## Calc daily cases from John Hopkins data
##
JHU <- dbGetQuery(db,"select * from JHUDATA where country='US'")
JHU$date <- as.Date(JHU$date)
CASES <- diff(JHU$confirmed,lag=1)
CASES <- as.data.frame(CASES)
CASES$YMD <- seq(as.Date("2020-01-23"),as.Date("2020-11-10"),"day")

```
```{r}
##
## Calc daily deaths
##
JHU <- subset(JHU,date >="2020-02-28")
DEATHS <- diff(JHU$deaths,lag=1)
DEATHS <- as.data.frame(DEATHS)
DEATHS$YMD <- seq(as.Date("2020-02-29"),as.Date("2020-11-10"),"day")
mean(CASES$CASES)

mean(DEATHS$DEATHS)
CASES$YMD <- as.character(CASES$YMD)
dbWriteTable(db, "JHCASES",CASES ,overwrite=TRUE)

DEATHS$YMD <- as.character(DEATHS$YMD)
dbWriteTable(db, "JHDEATHS",DEATHS ,overwrite=TRUE)
CASES$YMD <- as.Date(CASES$YMD)
DEATHS$YMD <-as.Date(DEATHS$YMD)
```

### US COVID19 Cases by Date
#### Source European CDPC

```{r}
ggplot(AB) + geom_line(aes(x=Reported,y=Cases, col="Daily Cases")) +  geom_smooth(aes(x=Reported,y=Cases,col="Loess"),span=0.25) +
  labs(title="European CDPC:  US daily Cases") +
  scale_x_date(date_breaks = "month")
```

#### Source John Hopkins University CSSE Cases

```{r}
ggplot(CASES) + geom_line(aes(x=YMD,y=CASES,col="Daily Cases")) +
    geom_smooth(aes(x=YMD,y=CASES),span=0.25) +
  labs(title="John Hopkins CSSE: US DAily Cases") +
  scale_x_date(date_breaks = "month")
```

### Source New York Times: Daily Cases

```{r}
ggplot(NYTCASES) + geom_line(aes(x=date,y=new_cases,col="Daily Cases")) +
   geom_smooth(aes(x=date,y=new_cases),span=0.25) +
  labs(title="New York Times: US Daily Cases") +
  scale_x_date(date_breaks = "month")
```

### Combined Three Sources:  US COVID19 Cases 

```{r}
plot1 <-ggplot() + geom_line(data=AB,aes(x=Reported,y=Cases, col="ECDPC")) +
   geom_line(data=CASES,aes(x=YMD,y=CASES,col="John Hopkins")) +
  geom_line(data=NYTCASES,aes(x=date,y=new_cases,col="NYTimes")) +
  labs(title="CombinedEuropean CDPC/NYTimes/John Hopkins: US Cases",y="Daily Cases") 
ggplotly(plot1)
  
```



### US COVID19 Deaths by Date

#### Source European CDPC

```{r}
ggplot(AB) + geom_line(aes(x=Reported,y=Deaths, col="Daily Deaths")) +  
  geom_smooth(aes(x=Reported,y=Deaths,col="Loess"),span=0.25) +
  labs(title="European CDPC:  US daily Deaths") +
  scale_x_date(date_breaks = "month")
```

#### Source John Hopkins University CSSE: US Deaths

```{r}
ggplot(DEATHS) + geom_line(aes(x=YMD,y=DEATHS,col="Daily Deaths")) +
  geom_smooth(aes(x=YMD,y=DEATHS),span=0.25) +
  labs(title="John Hopkins CSSE: US DAily Deaths") +
  scale_x_date(date_breaks = "month")
```

### Source New York Times: Daily Deaths

```{r}
ggplot(NYTDEATHS) + geom_line(aes(x=date,y=new_deaths,col="Daily Deaths")) +
    geom_smooth(aes(x=date,y=new_deaths),span=0.25) +
  labs(title="New York Times: US Daily Deaths") +
  scale_x_date(date_breaks = "month")
```

### Combined Three Sources:  US COVID19 Deaths 

```{r}
plot2 <-ggplot() + geom_line(data=AB,aes(x=Reported,y=Deaths, col="ECDPC")) +
   geom_line(data=DEATHS,aes(x=YMD,y=DEATHS,col="John Hopkins")) +
  geom_line(data=NYTDEATHS,aes(x=date,y=new_deaths,col="NYTimes")) +
  labs(title="CombinedEuropean CDPC/NYTimes/John Hopkins: US Deaths",y="Daily Deaths") 
ggplotly(plot2)
  
```

### Combined Tree Sources: Loess/Lowess Deaths

```{r}
plot2 <-ggplot() + geom_smooth(data=AB,aes(x=Reported,y=Deaths, col="ECDPC"),span=0.25,se=FALSE) +
   geom_smooth(data=DEATHS,aes(x=YMD,y=DEATHS,col="John Hopkins"),span=0.25,se=FALSE) +
  geom_smooth(data=NYTDEATHS,aes(x=date,y=new_deaths,col="NYTimes"),span=0.25,se=FALSE) +
  labs(title="CombinedEuropean CDPC/NYTimes/John Hopkins: US Deaths",y="Daily Deaths") 
ggplotly(plot2)
  
```

## exponential Moving Average


```{r}
# AA$EXP1<- movavg(AA$Cases,17,"e")
AA$EXP2<- movavg(AA$Cases,28,"e")
ggplot(AA) + 
  geom_line(aes(x=Reported,y=EXP2,col="28")) + 
  geom_line(aes(x=Reported,y=Cases,col="Cases"))
```

